\vspace{-2cm} \LARGE Divergence-Free Smoothed Particle Hydrodynamics \\ \large Advanced Simulation and Visualization of Fluids in Computer Graphics

### 2017-01-17 report/paper.tex
last changes

    Ihmsen et al.~\cite{ihmsen2014implicit} proposed an implicit incompressible SPH method with a pressure poisson equation made by the combination of a symmetric SPH pressure force and a SPH discretization of the continuity equation.
    The pressure poisson equation is used to compute the pressure so that the pressure forces correct the intermediate velocities to a divergence-free state.

    A predictive corrective incompressible SPH method was proposed by Solenthaler et al.~\cite{solenthaler} based on the Lagrangian SPH model.
    Incompressibility was enforced using a prediction correction scheme to determine the particle pressures.
    To achieve this the information about the density fluctuation was actively propagated through the fluid and the pressure values where updated until the targeted density was satisfied.

@@

    In the equations $\frac{D\rho}{Dt}$ and $\frac{D\textbf{v}}{Dt}$ denotes the derivate of the density $\rho$ and the velocity $\textbf{v}$ respectively.
    The pressure, kinematic viscosity and body forces are denoted by $p,v$ and $\textbf{f}$ respectively.
    The second part of Equation~\ref{eq:nsDensity}, $\nabla \cdot \textbf{v} = 0$, induce that the velocity field is free from divergence.
    From the continuity equation $\frac{D\rho}{Dt} = -\rho \nabla \cdot \textbf{v}$ and the divergence-free condition it follows that the partial derivate of the density with respect to the time is zero, which implies the equivalence of the equations in Equation~\ref{eq:nsDensity}.
    From this it follows that the density, in theory, must stay constant over time which is the implication of an incompressible fluid.
    In practice the divergence-free condition is not sufficient to guarantee incompressibility in simulations.
    The numerical time integrations will contain numerical errors which will cause density deviations due to the volume compressions.
    To avoid this a second condition $\rho - \rho_0 = 0$ called \textit{constant density condition} must be fullfilled~\cite{bender}.
    Section~\ref{densitySolver} and~\ref{divergenceSolver} give a deeper description of the solvers used in the project.

@@

    The stiffness parameters need to be determined iteratively since neighboring particles depend on each other.
    After the symmetric pressure force has been applied to the new velocity should the density change in Equation~\ref{eq:dpdt} be equal or close to zero.
    If the error is too large the process will be iterated until desired result have been achieved.

@@

All possible implicit geomitries can be descibed by a quadratic function, which is found in Equation~\ref{eqn:quadratic}, and in matrix form in Equation~\ref{eqn:quadricmatrixform}.

@@

    % Libraries
    % Computers
    % OpenGL
    % language
    % Makefiles
    The divergence-free SPH simulation was developed using OpenGL and C++ together with the OpenGL mathematics library GLM for vector calculations.
    FreeImage was used to save frames of the simulation and FFmpeg to assembled the frames into a video.
    PicoJSON was used to load in variables from a file making it easier to change variables quick and efficient and a makefile was used to compile the project.

    The simulation was developed on macOS and Linux.
    The Linux computer is running on Debian and has four Intel(R) Core(TM) i5-2450M CPU @ 2.50GHz, one NVIDIA GeForce 610M graphics card and 6.0 GiB RAM.
    One of the Mac OS computer is running on macOS Sierra and has a 1.8 GHz Intel Core i5 processor with an Intel HD Graphics 400 1536 MB graphics card.

@@

	\begin{figure}[h!]
		\includegraphics[width=\linewidth]{img/frame0001.png}
		\caption{Inital state of the water when the application starts.}
		\label{fig:res1}
	\end{figure}

>>>>

    Ihmsen et al.~\cite{ihmsen2014implicit} proposed an implicit incompressible SPH method with a pressure Poisson equation made by the combination of a symmetric SPH pressure force and a SPH discretization of the continuity equation.
    The pressure Poisson equation is used to compute the pressure so that the pressure forces correct the intermediate velocities to a divergence-free state.

    A predictive corrective incompressible SPH method was proposed by Solenthaler et al.~\cite{solenthaler} based on the Lagrangian SPH model.
    Incompressibility was enforced using a prediction correction scheme to determine the particle pressures.
    To achieve this the information about the density fluctuation was actively propagated through the fluid and the pressure values were updated until the targeted density was satisfied.

@@

    In the equations $\frac{D\rho}{Dt}$ and $\frac{D\textbf{v}}{Dt}$ denotes the derivate of the density $\rho$ and the velocity $\textbf{v}$ respectively.
    The pressure, kinematic viscosity and body forces are denoted by $p,v$ and $\textbf{f}$ respectively.
    The second part of Equation~\ref{eq:nsDensity}, $\nabla \cdot \textbf{v} = 0$, induce that the velocity field is free from divergence.
    From the continuity equation $\frac{D\rho}{Dt} = -\rho \nabla \cdot \textbf{v}$ and the divergence-free condition it follows that the partial derivate of the density with respect to the time is zero, which implies the equivalence of the equations in Equation~\ref{eq:nsDensity}.
    From this it follows that the density, in theory, must stay constant over time which is the implication of an incompressible fluid.
    In practice the divergence-free condition is not sufficient to guarantee incompressibility in simulations.
    The numerical time integrations will contain numerical errors which will cause density deviations due to the volume compressions.
    To avoid this a second condition $\rho - \rho_0 = 0$ called \textit{constant density condition} must be fulfilled~\cite{bender}.
    Section~\ref{densitySolver} and~\ref{divergenceSolver} give a deeper description of the solvers used in the project.

@@

    The stiffness parameters need to be determined iteratively since neighboring particles depend on each other.
    After the symmetric pressure force has been applied to the new velocity should the density change in Equation~\ref{eq:dpdt} be equal or close to zero.
    If the error is too large the process will be iterated until the desired result have been achieved.

@@

All possible implicit geometries can be described by a quadratic function, which is found in Equation~\ref{eqn:quadratic}, and in matrix form in Equation~\ref{eqn:quadricmatrixform}.

@@

    % Libraries
    % Computers
    % OpenGL
    % language
    % Makefiles
    The divergence-free SPH simulation was developed using OpenGL and C++ together with the OpenGL mathematics library GLM for vector calculations.
    FreeImage was used to save frames of the simulation and FFmpeg to assembled the frames into a video.
    PicoJSON was used to load variables from a file making it easier to change variables quickly and efficiently and a makefile was used to compile the project.

    The simulation was developed on macOS and Linux.
    The Linux computer is running on Debian and has a 50 GHz Intel Core i5-2450M processor and a NVIDIA GeForce 610M graphics card.
    One of the Mac OS computer is running on macOS Sierra and has a 1.8 GHz Intel Core i5 processor with an Intel HD Graphics 400 1536 MB graphics card.

@@

	\begin{figure}[h!]
		\includegraphics[width=\linewidth]{img/frame0001.png}
		\caption{Initial state of the water when the application starts.}
		\label{fig:res1}
	\end{figure}
